### 2021-12-20 papers/main.tex
TEMPORARY DEPLOY COMMIT

% \documentclass{uai2021} % for initial submission
\documentclass[]{uai2021} % after acceptance, for a revised
                                    % version; also before submission to
                                    % see how the non-anonymous paper
                                    % would look like
%% There is a class option to choose the math font
% \documentclass[mathfont=ptmx]{uai2021} % ptmx math instead of Computer
                                         % Modern (has noticable issues)
% \documentclass[mathfont=newtx]{uai2021} % newtx fonts (improves upon
                                          % ptmx; less tested, no support)
% NOTE: Only keep *one* line above as appropriate, as it will be replaced
%       automatically for papers to be published. Do not make any other
%       change above this note for an accepted version.

%% Choose your variant of English; be consistent
\usepackage[american]{babel}
% \usepackage[british]{babel}

%% Some suggested packages, as needed:
\usepackage{natbib} % has a nice set of citation styles and commands
    \bibliographystyle{plainnat}
    \renewcommand{\bibsection}{\subsubsection*{References}}
\usepackage{mathtools} % amsmath with fixes and additions
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
\usepackage{tikz} % nice language for creating drawings and diagrams
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{bm}

%% Provided macros
% \smaller: Because the class footnote size is essentially LaTeX's \small,
%           redefining \footnotesize, we provide the original \footnotesize
%           using this macro.
%           (Use only sparingly, e.g., in drawings, as it is quite small.)

%% Self-defined macros
\newcommand{\swap}[3][-]{#3#1#2} % just an example
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\state}{\mathcal{S}}
\newcommand{\action}{\mathcal{A}}
\newcommand{\KL}{\mathrm{KL}}

\title{Scalable and Parameter-Free Optimal Regret Reinforcement Learning}

% The standard author block has changed for UAI 2021 to provide
% more space for long author lists and allow for complex affiliations
%
% All author information is authomatically removed by the class for the
% anonymous submission version of your paper, so you can already add your
% information below.
%
% Add authors
\author[1]{\href{mailto:<jj@example.edu>?Subject=LALAL}{Author}{}}

\begin{document}
\maketitle

\begin{abstract}
%   This is the abstract for this article.
%   It should give a self-contained single-paragraph summary of the article's contents, including context, results, and conclusions.
%   Avoid citations; but if you do, you must give essentially the whole reference.
%   For example: This whole paper is devoted to praising É. Š. Åland von Vèreweg's most recent book (“Utopia's government formation problems during the last millenium”, Springevier Publishers, 2016).
%   Also, do not put mathematical notation and abbreviations in your abstract; be descriptive.
%   So not “we solve \(x^2+A xy+y^2\), where \(A\) is an RV”, but “we solve quadratic equations in two unknowns in which a single coefficient is a random variable”.
%   The reason is that mathematical notation will not display correctly when the abstract is reused on the proceedings website, for example, and that one should not assume the abstract's reader knows the abbreviation.
%   Of course the same remarks hold for your paper's title.
\end{abstract}

% \section{Introduction}
% Bayesian reinforcement learning is an approach to RL that expresses uncertainty in the
% Markov decision process (MDP) via a posterior distribution \citep{ghavamzadeh_bayesian_2015}.
% The posterior captures the uncertainty in transition and reward distributions.

\section{Background}
A reinforcement learning environement is modelled as a Markov decision process (MDP)
\(M = \langle \state, \action, r, P, P_0, \gamma \rangle\), where \(\state\) is the
state space and \(\action\) is the set of available actions. At time \(t=0\) a state
\(s_0\) is sampled from the distribution \(P_0(\cdot)\). At each timestep an action
\(a_t \sim \pi(\cdot \vert s_t)\) is selected and the agent transitions to a new state
state \(s_{t+1} \sim P(\cdot \vert s_t, a_t)\). A scalar reward
\(r_{t+1} = r(\cdot \vert s_{t+1}, s_t, a_t)\) is observed. As the agent and environement
interact in a sequence of time steps, a history of observations
\(\mathcal{H}_t = (s_0, a_0, s_1, r_1, a_1, s_2, r_2, \dots, s_t, r_t)\) is collected.
The goal is to find a policy \(\pi^\star\), such that sampling actions
\(a \sim \pi^*(\cdot \vert s)\) maximizes the expected accumulated and discounted future reward,
\(J^\pi \defeq \E_\pi \left[ \sum_{t=1}^\infty \gamma^t r_t \right]\). An efficient
agent must be able to learn from the data it collects, but since the data is
dependent on the policy, it must also prioritize to explore states and actions that
the agent can learn a lot from.

The efficiency of exploration can be measured in regret.
The regret of a policy is the difference in the expected reward obtained by
following that policy, and the expected reward for an optimal policy.
An learning algorithm's efficiency in exploration can be measured by its
cumulative regret over time. One of the simplest reinforcement learning 
problems is the multi-arm bandit problem. This is an MDP with no state and no
transition probabilities. One exploration strategy employed in most Q-learning
algorithms to date, \(\epsilon\)-greedy is provably inefficient, and has a
regret bound that grows linearly with time. There are several optimal algorithms
for this problem, but perhaps the simplest one is called Thompson sampling.
Thompson sampling approximates a posterior distribution of the expected reward
for each action. The next action is decided by sampling rewards for each action
from the posterior distribution, and selecting the action that gave the highest
reward. Bayesian methods been also been shown to behave efficiently with respect
to cumulative regret \cite{osband_generalization_2016} on general MDPs.
%\textcolor{red}{Optimal Regret + Thompson sampling }

The \(Q\)-function, \(Q^\pi(s,a)\), is defined as the expected reward of taking action
\(a\) in state \(s\) and then following policy \(\pi\) thereafter:
\(Q^\pi \defeq \E_\pi \left[ \sum_{t=0}^\infty \gamma^t r_t \vert s_0=S, a_0=A \right]\).
The Bellman operator on \(Q^\pi\) is defined as
\(\B[Q^\pi](s_t,a_t) \defeq \E_{P(s_{t+1} \vert s_t, a_t)\pi(a_{t+1} \vert s_t)}
\left[ r(s_{t+1}, s_t, a) + \gamma Q^\pi(s_{t+1}, a_{t+1}) \right]\). With the
\(Q\)-function we can define a policy that always picks the action with the 
highest \(Q\)-value in any state. We can't find \(Q\) exactly, but with a good
approximation this policy, \(\pi(a \vert s) = \arg\max_a Q(s,a)\), should
hopefully be close to optimal. In deep RL, we model \(Q\) with a
neural network \(\hat{Q}_\omega\). One way to facilitate exploration in this
policy is to introduce uncertainty in the
\(Q\)-value.
\citet{fortunato_noisy_2019} introduced NoisyNet for exploration.
These are networks with stochastic weights, where each weight \(w^{(l)}_{ij}\)
has an added pertrubaition sampled from a noise distribution with standard deviation \(\sigma^{(l)}_{ij}\).
After initializing \(\bm{w}\) and \(\bm{\sigma}\) such that the network has sufficient
stochasticity for exploration, both parameters are learned using standard backpropagation.
The approach is similar to variational inference schemes such as Bayes by backprop~\citep{blundell_weight_2015} where the weights of a neural network model are assumed to be normally distributed with mean 
\(\mu^{(l)}_{ij}\) and standard deviation \(\sigma^{(l)}_{ij}\). The objective of
Bayes by backprop, however, is to approximate the posterior distribution
\(p(\bm{w} \vert \mathcal{D})\) in a task with a fixed dataset \(\mathcal{D}\).
The parameter distribution in NoisyNet does not necessarily
converge to an approximate posterior. This means that it does not have the same
guarantee on total regret as methods that approximate a posterior over the
value functions~\citep{osband_generalization_2016}. They do, however, have
experimental results which shows that the function does not always converge to
a deterministic solution, but it is unclear why the network learns to introduce
more noise into the network parameters.

\citet{fortunato_noisy_2019} apply NoisyNet to three reinforcement learning
algorithms, DQN, Dueling DQN, and A3C, and show improved performance on all
of them. Later NoisyNet was used in the Rainbow 
algorithm~\citep{hessel_rainbow_2017}, a combination of six extensions to the
DQN algorithms~\citep{fortunato_noisy_2019, bellemare_distributional_2017, wang_dueling_2016,van_hasselt_deep_2015,schaul_prioritized_2016,mnih_asynchronous_2016}, that shows state-of-the-art performance across 57 Atari games.

%To properly adapt Bayesian inference to MDPs, \cite{fellows_bayesian_2021} define the Bayesian Bellman
%operator (BBO), \(P_B(b \vert s, a, \omega)\), as the distribution over
%Bellman functions such that for a noisy sample \(b_i \sim P_B(b \vert s,a,\omega)\),
%\(b_i = \B[\hat{Q}_\omega](s_i, a_i) + \eta_i\). Then approximates this with a
%distribution parameterized by \(\phi\) such that
%\(P(b \vert s,a,\phi) \approx P(b \vert s,a,\omega)\).
%\begin{equation}
%    \B_{\omega, N}^\star(s,a) \defeq \E_{P(\phi \vert \D_\omega^N)} \left[ \hat{B}_\phi(s,a) \right].
%\end{equation}
%Here, \(\hat{B}_\phi(s,a) = \E_{P(b \mid s, a, \phi)}[b]\), is the conditional mean of
%\(P(b \vert s, a, \phi)\), and the data \(\D_\omega^N\) is defined as the collection of
%sampled \(Q\)-values, states and actions \(\D_\omega^N \defeq \{b_i, s_i, a_i\}_{i=1:N}\).
%Under the assumption that each state \(s_i\) is drawn i.i.d from a distribution with
%support over \(\state\), or from an ergodic Markov chain defined over a \(\sigma\)-algebra that is
%countably generated from \(\state\) they find that \(\omega^\star\) such that
%\(\hat{Q}_{\omega^\star} = \B^\star_{\omega^\star, N}\) can be found 
%by minimising the mean squared Bayesian Bellman error (MSBBE):
%\begin{equation}
%    \mathrm{MSBBE}_N(\omega) \defeq \lVert \hat{Q}_{\omega} - \B^\star_{\omega, N} \lVert^2_{\rho, \pi},
%\end{equation}
%And that an unbiased estimate of \(\nabla_\omega \mathrm{MSBBE}_N(\omega)\) can be calculated
%as long as we can sample from \(P(\phi \vert \D_\omega^N)\).

%Nonlinear function approximators such as neural networks typically have an intractable
%posterior, so we cannot calculate the posterior analytically. Instead we use a posterior approximation
%\(q(\phi \vert \D_\omega^N) \approx P(\phi \vert \D_\omega^N)\), together with an algorithm
%for approximate inference. \citet{fellows_bayesian_2021} presents the Bayesian Bellman Actor-Critic
%using randomized priors \citep{osband_randomized_2018}, and show state of the art exploratory
%behaviour.

%For most approximate inference methods in deep learning, it is difficult to incorporate domain
%knowledge into the prior. Typically the prior distribution is defined on the parameters as \(P(\phi)\),
%while the knowledge exists in the function space \(P(a \vert s)\).
%This is also the case for randomized priors.
% We therefore want to

In NoisyNet \citep{fortunato_noisy_2019}, the Q-function space can be seen
as a stochastic process on the probability space \((\Omega, \mathcal{F}, P)\).
The stochastic process can simply be written as
\(\{Q(\bm{s}, \bm{a}) : (\bm{s}, \bm{a}) \in \state \times \action \}\). For
any \(\omega \in \Omega\), \(Q(\cdot, \cdot, \omega)\) is a sample function
mapping \(\state \times \action \rightarrow \mathbb{R}\). To simplify notation
we will denote sample functions as \(f\). NoisyNet, specifically, is a stochastic
process parameterized by its weights and their standard deviation. We will define
this parameterized function as \(g_\phi\), and their sampling distribution as
\(q_\phi \in \mathcal{Q}\). Given a noise vector \(\xi\), we have
\(f(\bm{s}, \bm{a}) = g_\phi(\bm{a}, \bm{s}, \xi)\).

Functional variational Bayesian neural networks \citep{sun_functional_2019} is 
a variational inference method for neural networks that approximates the 
posterior distribution in function space.
\citet{sun_functional_2019} show that for two stochastic processes \(P\) and \(Q\):

\begin{equation}
    \KL[P \Vert Q] = \sup_{n \in \N, \X \in \mathcal{X}^n} \KL \left[ P_\X \Vert Q_\X \right],
\end{equation}

where \(\X \in \mathcal{X}^n\) is a finite measurement set and \(P_\X, Q_\X\) are the marginal distributions
at \(\X\). They further show that if \(\f^\X\) are the function values at points \(\X\), then

\begin{multline}\label{eq:fkl}
    \nabla_\phi \KL[q_\phi(\f^X) \Vert p(\f^\X)] \\
    = \E \left[ \nabla_\phi \f^\X( \nabla_\f \log q(\f^\X) - \nabla_\f \log p(\f^\X)) \right].
\end{multline}

The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log q(\f^\X)\)
and \(\nabla_\f \log p(\f^\X)\). \(\nabla_\f \log q(\f^\X)\) is likely 
intractable, considering \(q_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f^\X)\)
can be easy to compute analytically. To reduce variance in the gradients we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log q(\f^\X)\), they use a spectral Stein gradient estimator
\citep{shi_spectral_2018}.

% \section{Background}
% \citet{fellows_bayesian_2021} defines the Bayesian Bellman operator (BBO) as:
% \begin{equation}
%     \B_{\omega, N}^\star(s,a) \defeq \E_{P(\phi \vert \D_\omega^N)} \left[ \hat{B}_\phi(s,a) \right],
% \end{equation}
% where \(\hat{B}_\phi(s,a) = \E_{P(b \mid s, a, \phi)}[b]\), is the conditional mean of the
% parametric distribution \(P(b \vert s, a, \phi)\), which is a model of the true
% data-generating distribution \(P(b \vert s, a, \omega)\). They further show that we can
% find \(\omega^\star\) such that \(\hat{Q}_{\omega^\star} = \B^\star_{\omega^\star, N}\)
% by minimising the mean squared Bayesian Bellman error (MSBBE):

% \begin{equation}
%     \mathrm{MSBBE}_N(\omega) \defeq \lVert \hat{Q}_{\omega} - \B^\star_{\omega, N} \lVert^2_{\rho, \pi}
% \end{equation}
% An unbiased estimate of \(\nabla_\omega \mathrm{MSBBE}_N(\omega)\) can be calculated as
% long as we can sample from \(P(\phi \vert \D_\omega^N)\).

% Nonlinear function approximators such as neural networks typically have an intractable
% posterior, so we cannot calculate the posterior analytically. Instead we use a posterior approximation
% \(q(\phi \vert \D_\omega^N) \approx P(\phi \vert \D_\omega^N)\), together with an algorithm
% for approximate inference. \citet{fellows_bayesian_2021} presents the Bayesian Bellman Actor-Critic
% using randomized priors \citep{osband_randomized_2018}, and show state of the art exploratory
% behaviour.

% For most approximate inference methods in deep learning, it is difficult to incorporate domain
% knowledge into the prior. Typically the prior distribution is defined on the parameters as \(P(\phi)\),
% while the knowledge exists in the function space \(P(a \vert s)\).

% Functional variational Bayesian neural networks \citep{sun_functional_2019} lets us define a prior on the
% \emph{function space}. In a Q-learning algorithm, this would be a function
% \(f : \state \times \action \rightarrow \R\). For now we generalize, and say that
% \(f : \mathcal{X} \rightarrow \mathcal{Y}\).
% \citet{sun_functional_2019} show that for two stochastic processes \(P\) and \(Q\):

% \begin{equation}
%     \KL[P \Vert Q] = \sup_{n \in \N, \X \in \mathcal{X}^n} \KL \left[ P_\X \Vert Q_\X \right],
% \end{equation}

% where \(\X \in \mathcal{X}^n\) is a finite measurement set and \(P_\X, Q_\X\) are the marginal distributions
% at \(\X\). They further show that if \(\f^\X\) are the function values at points \(\X\), then

% \begin{multline}\label{eq:fkl}
%     \nabla_\phi \KL[q_\phi(\f^X) \Vert p(\f^\X)] \\
%     = \E \left[ \nabla_\phi \f^\X( \nabla_\f \log q(\f^\X) - \nabla_\f \log p(\f^\X)) \right].
% \end{multline}

% The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log q(\f^\X)\) and \(\nabla_\f \log p(\f^\X)\).
% To estimate these log-density gradients, they use a spectral Stein gradient estimator \citep{shi_spectral_2018}.
% Given enough function value samples, the spectral Stein gradient estimator can estimate score functions
% for both in-distribution and out-of-distribution samples. This means that we can use it to estimate
% both gradients. \(\nabla_\f \log q(\f^\X)\) is likely intractable intractable, considering \(q_\phi\) is
% a neural network with stochastic weights. Depending on how we define the prior, however, \(\nabla_\f \log p(\f^\X)\)
% can be easy to compute analytically. To reduce variance in the gradients we aim to define tractable priors.

\section{Method}
% Motivation:
% * Optimal regret
% * Meaningful priors
We will present a method that allows the following:

\begin{itemize}
    \item Optimal regret \textcolor{red}{TODO: define optimal regret}
    \item Incorporate domain knowledge
\end{itemize}

\textcolor{red}{A natural step for finding
the optimal sampling distribution \(q_\phi\) is to minimize the KL-divergence
between it and the true Q-value function. This is different from the optimization
scheme in NoisyNet, which only tries to maximize the likelihood that samples from
\(q_\phi\) comes from the true Q-value function.} \textcolor{green}{This should probably go under method}

We would like to approximate the posterior distribution of the Q value function.
This would (1) give optimal regret, (2) let us incorporate domain knowledge
through an appropriate prior distribution.
We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(N\) samples
from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \KL \left[ q(\f^{\D_\omega^N}, \f^M) \Vert p(\f^{\D_\omega^N}, \f^M) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^M\) is \(f\) applied to
a set of i.i.d. random points \(M = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let

\begin{equation}
    M = \{(s_i, a_j) \; \forall s_i \in \D_\omega^N,\, \forall a_j \in \action\}. 
\end{equation}
%To justify this we need to show that \(supp(c) = \state \times \action\).
%Under the same assumption about the distribution of each state \(s_i\) from BBO,
%\(supp(c) = \state \times \action\) follows trivially.

Since we are modeling \(q_\phi\) with a Bayesian neural network, and \(Q_\omega\) with a neural network,
we have a bilevel optimization problem. To solve this, we employ a similar strategy to \citet{fellows_bayesian_2021},
where we have a two-timescale gradient update. \(\phi\) and \(\omega\) are updated using stochastic
gradient descent at different timescales to ensure stable convergence.

% We can now define a prior \(P(f)\), however, this prior exists in the space
% \(P(Q \vert s, a) : (\state \times \action \rightarrow \R) \rightarrow \R\), which can be hard to
% define. We would like the prior to be of the form \(P(a \vert s)\).

\begin{algorithm}
\caption{NoisyNet RL}
\begin{algorithmic}
    \State \(\D \gets \emptyset\)
    \State \(s \sim P_0\)
    \State Initialize \(\phi, \omega\)
    \While{not converged}
        \State \(f \sim q_\phi\)
        \State \(a \sim \arg\max_a f(s,a)\)
        \State \(s' \sim P(\cdot \vert s,a)\)
        \State \(r = r(s',a,s)\)
        \State \(\D \gets \D \cup \{s,a,r,s'\}\)
        \State \Call{UpdateNoisyNet}{$\phi, \omega, \D$}
    \EndWhile
    \Statex
    \Function{UpdateNoisyNet}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(f \sim q_\phi\)
        \State \(b \sim q_\omega\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q} \gets \max_a \f(s,a)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(G \gets r + \gamma b(s,a)\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert T \rvert} \nabla_\phi \log p(\bm{q} \vert \bm{G})\)
        \EndFor
        \State \(\phi \gets \phi + \eta_\phi \Delta_\mathcal{L}\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Functional Bayesian RL}
\begin{algorithmic}
    \State \(\D \gets \emptyset\)
    \State \(s \sim P_0\)
    \State Initialize \(\phi, \omega\)
    \While{not converged}
        \State \(f \sim q_\phi\)
        \State \(a \sim \arg\max_a f(s,a)\)
        \State \(s' \sim P(\cdot \vert s,a)\)
        \State \(r = r(s',a,s)\)
        \State \(\D \gets \D \cup \{s,a,r,s'\}\)
        \State \Call{UpdatePosterior}{$\phi, \omega, \D$}
    \EndWhile
    \Statex
    \Function{UpdatePosterior}{$\phi, \omega, \D$}
        \State \(T \sim \D\)
        \State \(f \gets \{f_i \sim q_\phi \mid i=1,\dots,k\}\)
        \State \(\mathbf{b} \gets \{b_i \sim q_\omega \mid i=1,\dots,k\}\)
        \State \(\mathcal{F} \gets \emptyset\)
        \State \(\Delta_\mathcal{L} \gets 0\)
        \For{\(\{s,a,r,s'\} \in T\)}
            \State \(\bm{q} \gets \max_a \f(s,a)\)
            \State \(\mathcal{F} \gets \mathcal{F} \cup \f(s,a)\)
            \State \(\bm{G} \gets \{r + \gamma b_i(s,a) \mid i = 1, 
        \dots, k\}\)
            \State \(\Delta_\mathcal{L} \gets \Delta_\mathcal{L} - \frac{1}{\lvert T \rvert} \nabla_\phi \log p(\bm{q} \vert \bm{G})\)
        \EndFor
        \State \(\Delta_\KL \gets SSGE(p, \mathcal{F})\)
        \State \(\phi \gets \phi + \eta_\phi (\Delta_\mathcal{L} - \lambda \Delta_\KL)\)
        \State \(\omega \gets \omega + \eta_\omega (\omega - \phi)\)
    \EndFunction
\end{algorithmic}
\end{algorithm}

\section{Experiments}
\subsection{Uninformative Prior}
\subsection{Informative Prior}

\section{Discussion}

\begin{acknowledgements} % will be removed in pdf for initial submission,
                         % so you can already fill it to test with the
                         % ‘accepted’ class option
    Briefly acknowledge people and organizations here.

    \emph{All} acknowledgements go in this section.
\end{acknowledgements}

\bibliography{uai2021-template}

\appendix
% NOTE: necessary when ptmx or no mathfont class option is given
% \providecommand{\upGamma}{\Gamma}
% \providecommand{\uppi}{\pi}
% \section{Math font exposition}
% How math looks in equations is important:
% \begin{equation*}
%   F_{\alpha,\beta}^\eta(z) = \upGamma(\tfrac{3}{2}) \prod_{\ell=1}^\infty\eta \frac{z^\ell}{\ell} + \frac{1}{2\uppi}\int_{-\infty}^z\alpha \sum_{k=1}^\infty x^{\beta k}\mathrm{d}x.
% \end{equation*}
% However, one should not ignore how well math mixes with text:
% The frobble function \(f\) transforms zabbies \(z\) into yannies \(y\).
% It is a polynomial \(f(z)=\alpha z + \beta z^2\), where \(-n<\alpha<\beta/n\leq\gamma\), with \(\gamma\) a positive real number.

\end{document}